Correcting for bias in high-throughput sequencing data

### 2011-09-25 seqbias-paper.tex
combined two figures

\begin{figure}
\centerline{\includegraphics[width=0.25\textwidth]{overview.eps}}
\caption{An overview of the approach taken: (a) foreground sequences are sampled
from the regions surrounding the starts of mapped reads, (b) background
sequences are sampled by randomly offsetting foreground positions, (c) a
Bayesian network is trained to discriminate between the set of
sampled foreground and background sequences, (d) and the model is evaluated
at each position within a locus, predicting bias.}
\label{fig:overview}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=0.4\textwidth]{uniformity.eps}}
\caption{Read counts (i.e., the number of reads starting at each position)
from the Mortazavi data set are plotted along the $3'$ UTR of Apoa2. The upper
plot shows unadjusted counts, the lower, counts adjusted by our method.}
\label{fig:unif}
\end{figure}

In this comparison, we measure how well the counts conform to a Poisson process.
The assumption of positional read counts following a Poisson distribution is
known to be a poor fit \citep{Srivastava2010}, but measuring the improvement in
the fit derived from correcting for bias remains a principled and easily
interpreted criterion. This increase in uniformity is illustrated in Figure
\ref{fig:unif}.

@@

We used sequencing data previously published by \citet{Au2010} to evaluate the
effect bias correction has on correlation to measurements made by TaqMan RT-PCR,
made available by the the Microarray Quality Control project \cite{Shi2006}. The
RNA-Seq data shows a pattern of bias similar to that seen in the other samples
sequenced on an Illumina platform.
%% TODO: refer to supplement for figure of bias

>>>>

\begin{figure}
\centerline{\includegraphics[width=0.4\textwidth]{overview.eps}}
\caption{An overview of the approach taken: (a) foreground sequences are sampled
from the regions surrounding the starts of mapped reads, (b) background
sequences are sampled by randomly offsetting foreground positions, (c) a
Bayesian network is trained to discriminate between the set of
sampled foreground and background sequences, (d) and the model is evaluated
at each position within a locus, predicting bias. The predicted bias can then be
used to adjust read counts, as in (e). In (d) and (e) we show the results of
this method applied to the $3'$ UTR of Apoa2, using data from
\citet{Mortazavi2008}.}
\label{fig:overview}
\end{figure}

@@

In this comparison, we measure how well the counts conform to a Poisson process.
The assumption of positional read counts following a Poisson distribution is
known to be a poor fit \citep{Srivastava2010}, but measuring the improvement in
the fit derived from correcting for bias remains a principled and easily
interpreted criterion. This increase in uniformity is illustrated in Figure
\ref{fig:overview}.

@@

We used sequencing data previously published by \citet{Au2010} to evaluate the
effect bias correction has on correlation to measurements made by TaqMan RT-PCR,
made available by the the Microarray Quality Control project \citep{Shi2006}. The
RNA-Seq data shows a pattern of bias similar to that seen in the other samples
sequenced on an Illumina platform.
%% TODO: refer to supplement for figure of bias
